---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
---

R Assignment #1
```{r}
clev = read.csv("ClevelandHeart.csv")
```
1. Amount of male patients:
```{r}
sum(clev$Sex == "male")
```
There are 206 male patients out of the 303 patients in the study.

2. The amount of entries missing from the Thal column:
```{r}
sum(is.na(clev$Thal))
```
2. The rows missing entries from the Thal column:
```{r}
which(is.na(clev$Thal))
```
2. All the data for these patients missing entries for the Thal column:
```{r}
clev[88,]
```
```{r}
clev[267,]
```
3. Scatterplot with Age on the x-axis and MaxHR on the y-axis:
```{r}
plot(clev$Age,clev$MaxHR,xlab="Age",ylab="Max HR",main="Age vs Max HR")
```
In the plot, you can see that as the age of the patients increased, their max heart rate tends to decrease.

4. Histogram 1
```{r}
hist(clev$Chol,xlab="Chol",main="Histogram 1")
```
4. Histogram 2:
```{r}
hist(clev$Chol,breaks=20,xlab="Chol",main="Histogram 2")
```
4. Boxplot:
```{r}
boxplot(clev$Chol,main="Chol Boxplot")
```
4. QQ Plot:
```{r}
qqnorm(clev$Chol,main="Chol Q-Q Plot")
```
Yes, there are outliers in the data set. These outliers are most apparent in the Q-Q plot and the box plot. In the box plot, 4-5 outliers are seen above the upper limit and in the Q-Q plot there are 5 points that are not in line with the rest of the data.
5. 
```{r}
mean(clev$Chol)
median(clev$Chol)
sd(clev$Chol)
IQR(clev$Chol)
```
6. Max Chol Value:
```{r}
max(clev$Chol)
```
Number of patients with normal Chol values:
```{r}
x = clev$Chol
sum(x<200)
```
Number of patients with high Chol values:
```{r}
sum(x>240)
```
7. Side by side boxplot of Chol values for men and women:
```{r}
boxplot(Chol~Sex,data=clev,main="Chol Values of Men and Women")
```
8. Histogram of male Chol values:
```{r}
mchol = clev[clev$Sex=="male",]
hist(mchol$Chol,main="Male Chol Histogram",xlab="Male Chol Values")
```
8. Male chol boxplot:
```{r}
boxplot(mchol$Chol,main="Male Chol Boxplot")
```
Male Chol Q-Q plot:
```{r}
qqnorm(mchol$Chol,main="Male Chol Q-Q Plot")
```
```{r}
mean(mchol$Chol)
median(mchol$Chol)
sd(mchol$Chol)
IQR(mchol$Chol)
```
Looking at the male chol data, it is hard to tell if there is an outlier from the histogram and boxplot. The most telling data is in the Q-Q plot, with a couple of data points being out of line with the rest towards the start of the graph.

9. Female chol histogram:
```{r}
fchol = clev[clev$Sex=="female",]
hist(fchol$Chol,main="Female Chol Histogram",xlab="Female Chol Values")
```
Female Chol Boxplot:
```{r}
boxplot(fchol$Chol,main="Female Chol Boxplot")
```
Female Chol Q-Q Plot:
```{r}
qqnorm(fchol$Chol,main="Female Chol Q-Q Plot")
```
```{r}
mean(fchol$Chol)
median(fchol$Chol)
sd(fchol$Chol)
IQR(fchol$Chol)
```
Looking at the female chol data, there is an outlier present shown in the boxplot at a much higher value than the upper fence. It is also present in the Q-Q plot as being out of line with the trend all the way in the corner of the graph.

10. Side by side boxplot of MaxHR for patients diagnosed and not diagnosed with heart disease:
```{r}
boxplot(MaxHR~AHD,data=clev,main="Max HR for Patients Diagnosed or Not Diagnosed with Heart Disease")

```

